---
title: "Case Study 10: Satellite Remote Sensing"
author: Wei Liu
date: November 10, 2021
output: github_document
---

## Load needed libraries

```{r, message=FALSE}
library(raster)
library(rasterVis)
library(rgdal)
library(ggmap)
library(tidyverse)
library(knitr)
library(ncdf4) # to import data from netcdf format
```

## Prepare the data

### Download the data

```{r, message=FALSE}
# Create afolder to hold the downloaded data
dir.create("data",showWarnings = F) #create a folder to hold the data

lulc_url="https://github.com/adammwilson/DataScienceData/blob/master/inst/extdata/appeears/MCD12Q1.051_aid0001.nc?raw=true"
lst_url="https://github.com/adammwilson/DataScienceData/blob/master/inst/extdata/appeears/MOD11A2.006_aid0001.nc?raw=true"

# download them
download.file(lulc_url,destfile="data/MCD12Q1.051_aid0001.nc", mode="wb")
download.file(lst_url,destfile="data/MOD11A2.006_aid0001.nc", mode="wb")
```

### Load data into R

```{r, message=FALSE, warning=FALSE}
lulc=stack("data/MCD12Q1.051_aid0001.nc",varname="Land_Cover_Type_1")
lst=stack("data/MOD11A2.006_aid0001.nc",varname="LST_Day_1km")
```

### Explore LULC data

```{r, fig.align='center'}
plot(lulc)

# pick one year to work with
lulc=lulc[[13]]
plot(lulc)
```

### Process land cover data

```{r, message=FALSE}
Land_Cover_Type_1 = c(
  Water = 0, 
  `Evergreen Needleleaf forest` = 1, 
  `Evergreen Broadleaf forest` = 2,
  `Deciduous Needleleaf forest` = 3, 
  `Deciduous Broadleaf forest` = 4,
  `Mixed forest` = 5, 
  `Closed shrublands` = 6,
  `Open shrublands` = 7,
  `Woody savannas` = 8, 
  Savannas = 9,
  Grasslands = 10,
  `Permanent wetlands` = 11, 
  Croplands = 12,
  `Urban & built-up` = 13,
  `Cropland/Natural vegetation mosaic` = 14, 
  `Snow & ice` = 15,
  `Barren/Sparsely vegetated` = 16, 
  Unclassified = 254,
  NoDataFill = 255)
```

```{r}
lcd=data.frame(
  ID=Land_Cover_Type_1,
  landcover=names(Land_Cover_Type_1),
  col=c("#000080","#008000","#00FF00", "#99CC00","#99FF99", "#339966", "#993366", "#FFCC99", "#CCFFCC", "#FFCC00", "#FF9900", "#006699", "#FFFF00", "#FF0000", "#999966", "#FFFFFF", "#808080", "#000000", "#000000"),
  stringsAsFactors = F)
# colors from https://lpdaac.usgs.gov/about/news_archive/modisterra_land_cover_types_yearly_l3_global_005deg_cmg_mod12c1
kable(head(lcd))
```

```{r, message=FALSE, results='hide'}
# convert to raster (easy)
lulc=as.factor(lulc)
# update the RAT with a left join
levels(lulc)=left_join(levels(lulc)[[1]],lcd)
```

### Plot LULC data

```{r}
gplot(lulc)+
  geom_raster(aes(fill=as.factor(value)))+
  scale_fill_manual(values=levels(lulc)[[1]]$col,
                    labels=levels(lulc)[[1]]$landcover,
                    name="Landcover Type")+
  coord_equal()+
  theme(legend.position = "bottom")+
  guides(fill=guide_legend(ncol=5, nrow=3, byrow=TRUE))
```

### Plot land surface temperature

```{r}
plot(lst[[1:12]])
```

### Convert LST to Degrees C

```{r}
offs(lst)=-273.15
plot(lst[[1:10]])
```

### Add Dates to Z (time) dimension

```{r}
names(lst)[1:5]

tdates=names(lst)%>%
  sub(pattern="X",replacement="")%>%
  as.Date("%Y.%m.%d")

names(lst)=1:nlayers(lst)
lst=setZ(lst,tdates)
```

## Part 1: Extract timeseries for a point

Extract LST values for a single point and plot them.

```{r, message=FALSE, warning=FALSE}
lw=SpatialPoints(data.frame(x= -78.791547,y=43.007211))
projection(lw) <- "+proj=longlat"
lw_transformed = spTransform(lw, crs(lst))
lst_data = raster::extract(lst,lw_transformed,buffer=1000,fun=mean,na.rm=T)
lst_data_transposed = t(lst_data)
dates = getZ(lst)
dates_lst = data.frame(date = dates, lst = lst_data_transposed)

ggplot(dates_lst, aes(date, lst)) +
  geom_point() +
  geom_smooth(span=0.01, n=100) +
  labs(x = "Time",
       y = "Monthly Mean Land Surface Temperature")
```

## Part 2: Summarize weekly data to monthly climatologies

```{r}
tmonth = as.numeric(format(getZ(lst),"%m"))
lst_month = stackApply(lst, tmonth, fun=mean)
names(lst_month)=month.name
```

Then we plot the map for each month.

```{r}
gplot(lst_month) + 
  geom_tile(aes(fill = value)) +
  scale_fill_gradient(low = 'dark blue', high = 'orange') +
  facet_wrap(~variable) +
  coord_equal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.3))
```

Calculate the monthly mean for the entire image.

```{r}
cellStats(lst_month,mean) %>% 
  as.data.frame() %>%  # Inspired by Hui
  rename(Mean = ".") %>% 
  kable()
```

